Adaptive Hybrid Screening for Efficient Lasso Optimization

### 2021-09-11 jscs/response/response.tex

Thank you very much for your detailed reading of the paper and your insightful comments and questions. You raise a number of good points and we have modified the manuscript in several places based upon your feedback.  In our opinion, these revisions have nicely improved upon the original submission.

\begin{enumerate}

\item \emph{Numerical studies only focus on the computational performances of the algorithms. One may be interested in some details in the proposed algorithm: the size of variables remaining from the safe or strong rules, the number of KKT condition checks, the adaptive batch sizes, etc. It would be good to add the numerical studies for details in the algorithm.}

  This is an excellent point. We felt that some readers would definitely want to know this additional information, while others might just care that the algorithm is faster. So, in the revised manuscript we present these additional details in an appendix, which does indeed lend insight into how all of the algorithms are working and how the proposed method manages to achieve efficiency gains.

@@

\item \emph{The paper assumes that the predictive variables are standardized for linear and logistic regression models. For group lasso, the additional orthogonal assumption are required for the proposed method. I’m just wondering if the proposed method is applicable without the standardization assumptions. Specifically, is it possible to develop (or exist) the sequential safe and strong rules without the standardization assumptions.}

  Most sequential safe or strong rules do not require standardization, including the rules that we describe in our manuscript. As a result, our method does not require standardization. We introduce the standardization assumption mainly for simplicity. In the revised manuscript, we have clarified this point at the end of the first paragraph in Section 2.

@@

\item \emph{In page 8, line 19, ``the reference solutions are continually updated''. The reference solutions are not always updated. So, I’m not sure the term ``continually'' is appropriate.}

  Very good point. In the revised manuscript, we have changed this to ``the reference gets updated systematically'', which is a more accurate description.

\item \emph{In Figure 3, the green line (HSSR) pattern is quite different with others. Could you explain it?}

  We have added a short explanation in Section 5.1.1 that HSSR works better when $\lambda$ values in the path are closer to $\lambda_{\max}$. Further exploration and discussion of the performance of hybrid rules compared to safe/strong rules may be found in Zeng et al.~(2021) paper in which they are proposed.

@@

\item \emph{In Section 6(Conclusion), there does not exist the sequential safe rules for sparse Cox regression, Poisson regression, and support vector machines. It would be good to explain the existence of the sequential strong rules for each problem.}

  Thank you for this good suggestion. In the final paragraph of the conclusion section, we now discuss the existence of strong rules for various models.

@@

\item \emph{In Theorem 4.1, $x_* \in \{ x_j : \hat{\beta}_j(\lambda_l) \ne 0 \}$. Is the $x_*$ arbitrary? It seems that further explanation for $x_*$ is needed.}

  This is a good question. We have added further clarification of this point in paragraph following the theorem. The feature $x_*$ is indeed arbitrary -- any feature with non-zero coefficient at the reference $\lambda_l$ suffices.

@@

\subsubsection*{Minor corrections:}

\begin{enumerate}

\item \emph{In Theorem 4.1, $\lambda_0 := \max_j |x^T_j y/n|$ may be incorrect. It would be the maximum gradient value at $\hat{\beta} = 0$ of the negative log-likelihood in logistic regression model.}

  You are correct that $\lambda_0$ should be the maximum gradient of the negative log-likelihood. However, under the standardization assumption, this gradient is $|x^T_j y/n|$ at $\hat{\beta} = 0$. Thus, the expression is correct, albeit not immediately obvious.

@@

\item \emph{In page 10, line 20, $|S_{l=b}|$ should be $|S_{l+b}|$.}

  Thank you very much for pointing this out. We have corrected this in the revised submission.

@@

\item \emph{In Figure 1, the green line is not continuous. It would be better to modify Figure 1.}

  Thank you for directing our attention to the fact that this may be confusing to the reader. The discontinuities are correct -- every time the reference is updated, this causes a ``jump'' as the screening power is greatly increased every time this happens. In the revised manuscript, we have added an explanation for this phenomenon in the paragraph following Figure 1.

@@

\item \emph{In page 9, line 50, the ’line 12’ is better than ’line 11’ in Algorithm 1.}

  Thank you very much for pointing this out. We have adopted this suggestion in the revised submission.

### 2021-09-11 jscs/adaptive-hssr.tex
Some edits to responses; new cover letter

This complexity analysis will only consider calculations with $O(np)$ complexity. The remaining calculations, those with $O(p)$ complexity for example, are not only of lower order, but also consume less memory and thus usually do no require out-of-core computing. Calculation of $\{x_j^Ty\}_{j=1}^p$ is needed only once for the whole solution path; we ignore this as well as its computational cost is the same across all approaches. We assume that strong rule screening is always powerful in the sense that its set of non-discarded features, $\mathcal{R}_{l+b}$, satisfies $|\mathcal{R}_{l+b}|=o(p) \approx 0$, but that safe rule screening, with corresponding set $\mathcal{S}_{l+b}$, may not be. We also assume that post-convergence check will usually pass in the first iteration and thus is only needed approximately once per $\lambda$. These two assumptions are supported by experiments from the origin paper of SSR \citep{Tibshirani2012}, many other papers studying feature screening, and our manuscript as discussed in the Appendix.

@@

When computing the adaptive Slores, $\{x_j^Ty\}_{j=1}^p$ needs to be computed only once, after which it can be stored and reused for the whole solution path. The vector $x_*$ can be chosen arbitrarily from features that has non-zero coefficients at the reference $\lambda_l$, for example, the feature whose coefficient that first enters the model. The quantity $\{x_j^Tx_*\}_{j=1}^p$ can be similarly reused unless the corresponding regression coefficient drops out of the model; this can happen but is relatively rare. Other than those two terms, the only term that requires expensive $O(np)$ computations is $\{x_j^Tr(\lambda_l)\}_{j=1}^p$, which needs to be computed only when we update the reference solution. Thus Algorithm~\ref{alg:path} can be applies to the sparse logistic model, with Slores replacing EDPP, to form the adaptive hybrid rule algorithm for sparse logistic regression.

@@

AHR again outperforms other screening methods on all data sets, faster than other methods by factor ranging from 70\% to 630\%. Also note that the performance of SEDPP varies considerably for the MNIST and NYT data sets, where the response is randomly selected in each replication, indicating that the speed of SEDPP is sensitive to the data. AHR is built with SEDPP as the safe rule but does not share this issue of instability because it can adaptively adjust itself according the performance of the safe rule. Besides the computation time, we also look at a variaty of other statistics in the experiments, as discussed in the Appendix.

@@

>>>>

This complexity analysis will only consider calculations with $O(np)$ complexity. The remaining calculations, those with $O(p)$ complexity for example, are not only of lower order, but also consume less memory and thus usually do no require out-of-core computing. Calculation of $\{x_j^Ty\}_{j=1}^p$ is needed only once for the whole solution path; we ignore this as well as its computational cost is the same across all approaches. We assume that strong rule screening is always powerful in the sense that its set of non-discarded features, $\mathcal{R}_{l+b}$, satisfies $|\mathcal{R}_{l+b}|=o(p) \approx 0$, but that safe rule screening, with corresponding set $\mathcal{S}_{l+b}$, may not be. We also assume that post-convergence check will usually pass in the first iteration and thus is only needed approximately once per $\lambda$. These two assumptions are supported by experiments from the original SSR paper \citep{Tibshirani2012} as well as our own investigations as discussed in the Appendix.

@@

When computing the adaptive Slores, $\{x_j^Ty\}_{j=1}^p$ needs to be computed only once, after which it can be stored and reused for the whole solution path. The vector $x_*$ can be chosen arbitrarily -- any feature with non-zero coefficient at the reference $\lambda_l$ suffices. The quantity $\{x_j^Tx_*\}_{j=1}^p$ can be similarly reused unless the corresponding regression coefficient drops out of the model; this can happen but is relatively rare. Other than those two terms, the only term that requires expensive $O(np)$ computations is $\{x_j^Tr(\lambda_l)\}_{j=1}^p$, which needs to be computed only when we update the reference solution. Thus Algorithm~\ref{alg:path} can be applies to the sparse logistic model, with Slores replacing EDPP, to form the adaptive hybrid rule algorithm for sparse logistic regression.

@@

AHR again outperforms other screening methods on all data sets, faster than other methods by a factor ranging from 70\% to 630\%. Also note that the performance of SEDPP varies considerably for the MNIST and NYT data sets, where the response is randomly selected in each replication, indicating that the speed of SEDPP is sensitive to the data. AHR is built with SEDPP as the safe rule but does not share this issue of instability because it adaptively adjusts itself according the performance of the safe rule. A more detailed investigation of the algorithms' performance on these data sets including information on the number of rejections and KKT checks required is provided in the Appendix.

@@

%\appendix
%\appendixpage
